Time Series Analysis of Chaotic Dynamical Systems with Recurrent Neural Networks

### 2019-04-18 doc/paper/paper.tex
Updates from Overleaf

Figure \ref{fig:doub_pend_diff} shows a plot of each of the dataset's
dimensions. It seems like the dataset's activity is fairly sparse. There is
only gradual but seemingly random variations until a few spots around $t=10,
30, 60, \text{and} 75$. The results for the ESN weren't very strong, but it
was able to predict almost two Lyapunov Times before diverging.

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_0_param_77_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_0_param_225_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_1_param_63_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_1_param_230_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_2_param_305_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_2_param_458_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_3_param_52_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_3_param_151_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_4_param_127_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_4_param_300_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

@@

The case I analyze, the initial conditions were found to be 
$(x, x', y, y')_{t=0} = (0.79563053, 0, 0, 0.96833226)$
with a largest Lyapunov Exponent of 0.1759. This case is much different than 
the other systems since it doesn't have strong chaotic properties. This can
be seen in the orbital pattern which seems quasi-periodic, and it might be. 
In light of this, it's also very possible that the measured chaotic properties
are solely due to error in the numerical calculation. However, the orbiting
doesn't repeat exactly, so there are still small complexities for the ESN
to pick up on. 
%% CONCLUSION %%%
\section{Conclusion}

>>>>

Figure \ref{fig:doub_pend_diff} shows a plot of each of the dataset's
dimensions. It seems like the dataset's activity is fairly sparse. There is
only gradual but seemingly random variations until a few spots around $t=10,
30, 60, \text{and} 75$. The results for the ESN weren't very strong, and none 
could predict more than one Lyapunov Times before diverging.

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_0_param_77_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_0_param_77_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_1_param_63_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_2_param_305_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_2_param_305_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_2_param_305_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_3_param_52_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_3_param_52_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_4_param_127_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_4_param_127_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

@@

The case I analyze, the initial conditions were found to be 
$(x, x', y, y')_{t=0} = (0.79563053, 0, 0, 0.96833226)$
with a largest Lyapunov Exponent of 0.1759. This case is much different than 
the other systems since it doesn't have strong chaotic properties. This can
be seen in the orbital pattern which seems quasi-periodic, and it might be. 
In light of this, it's also very possible that the measured chaotic properties
are solely due to error in the numerical calculation. However, the orbiting
doesn't repeat exactly, so there are still small complexities for the ESN
to pick up on. In the results, we see that the ESN performs very well, but
it's possible that the ESN is just converging to the mean of the orbits. 
The steep increase in error and the subsequent slowing of increase in error
seen in the graphs below suggests that this might be the case.
Since the orbits are fairly close, this results in a sufficiently
accurate approximation.

Most RMSE graphs look very similar. There is a steep increase in RMSE at first
but then it slows. Overall, the RMSE is really strong. Most don't rise above 
0.2, and one show below (Figure \ref{fig:3brounded}) shows an RMSE not rising 
above 0.1 despite its not as elliptical looking shape. Further, this model
is impressive because it only has 5 internal units. Most other top performers 
have over 100.

\begin{figure}[H]
    \centering
    \includegraphics[width=.5\linewidth]{doc/paper/images/r3body/full_differential.png}
    \caption{The entire difference dataset}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.5\linewidth]{doc/paper/images/r3body/test_data.png}
    \caption{The testing dataset partition}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_3_param_15_fit.png}
        \caption{Predicted Trajectory}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_3_param_15_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{The best validation and testing performance and only 5 internal units}
    \label{fig:3brounded}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_0_param_566_fit.png}
        \caption{Predicted Trajectory}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_0_param_566_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Best performer on validation set}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_1_param_366_fit.png}
        \caption{Predicted Trajectory}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_1_param_366_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{A top performer on validation set}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_2_param_262_fit.png}
        \caption{Predicted Trajectory}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_2_param_262_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{A top performer on validation set}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_4_param_448_fit.png}
        \caption{Predicted Trajectory}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/r3body/rank_4_param_448_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{A top performer on validation set}
\end{figure}

%% CONCLUSION %%%
\section{Conclusion}
